---
title: "Usage of fars functions"
author: "Giulia Pais"
date: "27-06-2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fars)
library(dplyr)
library(readr)
```
# Using fars functions to import, summarize and visualize Fatality Analysis Reporting System data
The fars package includes a variety of functions that allow you to analyze files from the Fatality Analysis
Reporting System. Let's go through a basic workflow, how should you use this package?

### If you're interested in import only
For importing one or more files to analyze you can use two different functions, `fars_read` or `fars_read_years`.
`fars_read` takes only one string object as a parameter which represents the path to the file you want to import
and it's most useful when you want to import a single dataset:
```{r echo=FALSE, eval=TRUE}
pathToFile <- system.file("extdata","accident_2013.csv.bz2",package ="fars")
```
```{r echo = TRUE}
df <- fars_read(pathToFile)
head(df)
```
`fars_read_years` takes a vector of positive integers as input, which represents the years that the datasets are
referring to. Please note that to be able to import the datasets with year only, the csv files should be in your
default directory or current directory if you have a project open. If you'd like to change your current directory you shoud use `#setwd("appropriate/path")`.

```{r echo = TRUE}
new <- system.file("extdata",package ="fars")
setwd(new)
dfy <- fars_read_years(c(2013,2014,2015))
head(dfy)
```

A list of tibbles is returned, each of them showing only month and year columns.

### Importing and summarizing
If you're interested in summarizing the data counting the number of accidents for year and month you should use
`fars_summarize_years`: as previously, this function takes a vector of positive integers representing years as a vector. It relies internally on `fars_read_years`, so all of the previous recommendations are still valid.
```{r echo = TRUE}
new <- system.file("extdata",package ="fars")
setwd(new)
summary <- fars_summarize_years(c(2013,2014,2015))
head(summary)
```

### Visualizing the accidents on map
It's useful to have a visualization of all the accidents that occurred in a single state and in a certain year.
To do this simply use `fars_map_state`:
```{r echo = TRUE}
new <- system.file("extdata",package ="fars")
setwd(new)
fars_map_state(1, 2013)
```
By providing a valid state number and a year a map is plotted displaying the coordinates of the various accidents happened in 2013.
